---
title: "South Australian COVID-19 Data Analyses"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: spacelab
---

```{r setup, include=FALSE}
library(flexdashboard)

library(janitor)
library(tidyverse)
library(lubridate)
library(tidyquant)
library(googlesheets4)
library(tidyquant)
library(ggrepel)
library(plotly)
library(DT)
library(tibbletime)

# gather the data
data_raw_tbl <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1AHOf4y6mkvizSMG2GZ1GSHFRHmi_PteCX9r8fAmXhlk/edit?usp=sharing", sheet="Data")

# get the date for labelling
date_label <- format(now(), "%d-%m-%Y")


```

Overview
===================================

Column
-----------------------------------
 

### SA Health Covid Data - Transcribed from SA Health press releases

This dashboard presents some simple analyses of data collated from press releases related to the South Australian COVID-19 outbreak made by the SA Department of Health. 

Why? Well I live in Adelaide and data was not being presented at the State level at the time I started this, only via press releases. So I started transcribing the data and doing some charts. 

I'm particularly interested in the rates and log-log behaviour to the only way to do that was collect my own data and prepare plots 

The SA Government has lots of really nice data available at their dashboard page https://www.covid-19.sa.gov.au/home/dashboard so please check that out as it's the official data 

__Any errors are due to me, this is not official data__

The process - 

* Data are transcribed from the SA Health Covid press releases on the day they are released, or as close to it as possible from this link https://www.sahealth.sa.gov.au/wps/wcm/connect/public+content/sa+health+internet/about+us/news+and+media/all+media+releases/all+media+releases
* The data are entered into a spreadsheet on Google Sheets at the following link https://docs.google.com/spreadsheets/d/1AHOf4y6mkvizSMG2GZ1GSHFRHmi_PteCX9r8fAmXhlk/edit?usp=sharing
* An R script is used to tidy and analyse the data and then prepare this dashboard using the `flexdashboard` library

Analyses so far -

* Time Series Plots
* Total vs Daily New Cases 
* Daily Growth Factor vs Time
* Percentage Deaths per Case

Click on the `Charts` tab and then select the appropriate tab title. You can also click on the `Data` tab to view the raw data

Dashboard last updated 30th April 2020

Code here https://github.com/morebento/covid_sa

Charts
===================================


Row {.tabset .tabset-fade}
------------------------------------

### Time Series Plots

Active Cases is calculated as `Total Cases - Cleared of Covid Total - Deaths Total`

```{r}

time_series_plot <- data_raw_tbl %>%
    select(Date, `Total Cases`, `Daily New Cases`, `Cleared of Covid Total`, `Deaths Total`) %>%
    mutate(
        `Active Cases` = `Total Cases` - `Cleared of Covid Total` - `Deaths Total`
    ) %>%
    gather(attribute, value, -Date) %>%
    drop_na() %>%
    clean_names() %>%
    ggplot(aes(date, value)) +
    geom_line(aes(colour=attribute)) +
    scale_colour_tq() +
    theme_tq() +
    labs(
        title = "Total, Daily and Cleared Cases vs Time",
        caption = str_glue("{date_label}, @morebento"),
        colour = "Measure"
    )

ggplotly(time_series_plot)

```


### Total vs Daily New Cases 

Inspried by Minute Physics' video https://www.youtube.com/watch?v=54XLXg4fYsc "Exponential increase on this chart should look like a line with a constant slope of +1.0.\" So anything dropping off to the right represents the outbreak is being bought under control. Axes are Log 10 transformed"


```{r}
log_plot <- data_raw_tbl %>%
    ggplot(aes(`Total Cases`, `Daily New Cases`, label=Date)) +
    geom_point(colour=palette_light()[1]) +
    #geom_ma(n=3) +
    geom_smooth(se=FALSE, aes(colour="Three Point Moving Average")) +
    scale_x_log10() +
    scale_y_log10() +
    scale_colour_tq() +
    theme_tq() +
    labs(
        title  = "Total vs Daily New Cases ",
        caption =  str_glue("{date_label}, @morebento")
    )

ggplotly(log_plot)

```

### Daily Growth Factor vs Time

5 day rolling mean and naive (previous day) rates provided. Inspired by \"The one COVID-19 number to watch\". https://www.abc.net.au/news/2020-04-10/coronavirus-data-australia-growth-factor-covid-19"

```{r}


# define a rolling mean using rollify
rolling_mean <- rollify(mean, window=5)

growth_factor_plot <- data_raw_tbl %>%
    clean_names() %>%
    select(date, total_cases) %>%
    mutate(
        previous_total_cases = lag(total_cases, order_by = date)
    ) %>%
    mutate(
        rolling_mean_5 = rolling_mean(total_cases)
    ) %>%
    mutate(
        rate_5_day_rolling_mean = total_cases / rolling_mean_5,
        rate_naive = total_cases / previous_total_cases
    ) %>%
    select(date, rate_5_day_rolling_mean, rate_naive) %>%
    gather(metric, value, -date) %>%
    drop_na() %>%
    mutate(
        metric = str_replace_all(metric, "_", " "),
        metric = str_to_title(metric)
    ) %>%
    ggplot(aes(date, value)) +
    geom_line(aes(colour=metric)) +
    #ylim(0.75, 1.75) +
    scale_colour_tq() +
    theme_tq()  +
    labs(
        title = "Daily Growth Factor vs Time",
        y = "Daily Growth Factor",       
        x = "Date",
        caption = str_glue("{date_label}, @morebento"),
        color = "Metric"
    )


ggplotly(growth_factor_plot)

```

### Percentage Deaths per Case 

Deaths per case as a percentge is calculated and then plotted with a three point moving average. 

```{r}

death_plot <- data_raw_tbl %>%
    clean_names() %>%
    select(date, total_cases, deaths_total) %>%
    mutate(
        deaths_per_case_pc = (deaths_total / total_cases) * 100
    ) %>%
    ggplot(aes(date, deaths_per_case_pc)) +
    geom_point(aes(colour="Actual Data")) +
    geom_smooth(se=FALSE, aes(colour="Three Point Moving Average")) +
    ylim(0, 1) +
    scale_colour_tq() +
    theme_tq()  +
    labs(
        title = "Percentage Deaths per Case ",
        y = "Deaths per Case (%)",       
        x = "Date",
        caption = str_glue("{date_label}, @morebento"),
        colour="Data Type"
    )
    

ggplotly(death_plot)

```


Data
===================================


Row {data-height=200}
------------------------------------


### Transcribed Data

```{r}
DT::datatable(data_raw_tbl)
```